---
title: "Lab work 4."
author: "Anatoly Makarevich"
date: '2018-12-13'
output: html_document
bibliography: bibliography.bib

---

```{r setup, include=FALSE}
library(crayon)
#library(xtable)
library(knitr)
library(moments)

library(arm)  # pretty histogram plot :)

library(fitdistrplus)  # 
```

```{r , echo=TRUE}
Nvar = 30 %% 24  # == 6
```

### Exercise 1

Generate 200 elements from (Var 6) *geometric(0.8)*. 

Output the following:

- first 10 and last 20.

- min, max

- mean, sd, median, skewness(assymetry), kurtosis, 25th/75th/95th percentiles


```{r , echo=TRUE}
# Geom (0.8)
sample <- rgeom(200, 0.8)

P <- function(txt, val, round_to=4) {
  val2 <- round(val, round_to)
  print(txt %+% as.character(val2))
}

show_it <- function(sample) {
  print("Head and tail:")
  print(sample[1:10])
  print(sample[(length(sample) - 20) : length(sample)])
  
  P("Min: ", min(sample))
  P("Max: ", max(sample))
  P("Median: ", median(sample))
  P("Mean: ", mean(sample))
  P("SD: ", sd(sample))
  print("Quantiles:")
  quantile(sample, probs = c(0.25, 0.75, 0.95))
  P("Skewness: ", skewness(sample))
  P("Kurtosis: ", kurtosis(sample))
}
```


Using an **alternate** method to print (it's a lot cleaner! its\_so\_awesome).

```{r, echo=FALSE}
p=0.8
print(
  c(
    mean=(1-p)/p, 
    sd=(1-p)^0.5/p, 
    median=qgeom(0.5, p), 
    skew=(2-p)/(1-p)^0.5, 
    kurtosis=6+p^2/(1-p),
    quantile=qgeom(c("25%"=0.25,"75%"=0.75,"95%"=0.95), p)
  )
)

```


Some of the formulas used:

$$ \mu = \frac{1}{p}; \quad 
\hat{\mu}= \frac{1}{n}  \sum \limits _{i=1} ^{n} x_{i}; \quad

\kappa = 6 + \frac{p^2}{\sqrt{1-p}}
$$


##### Boxplot:

```{r, echo=FALSE}
boxplot(sample)
title("Water Dropping Onto a Plate (Boxplot)")
```



```{r}
plot.ecdf(sample, main="Archery Practice (ECDF)")
```

#### Histogram, theoretical and actual

```{r, echo=FALSE}
x <- min(sample):max(sample)
y <- dgeom(x, p=0.8)

discrete.histogram(x, table(sample), y, xlab="sample (blue), theoretical (red)")
title("Actually looks pretty good!")
```

***


### Exercise 2

Using the Kolmogorov–Smirnov test doesn't work as we have a discrete distribution. 

```{r, echo=FALSE}
ks.test(sample, "pgeom", p=0.8, alternative = "two.sided")
```

Chi-squared (Pearson) test:

```{r, include=FALSE}

test_geom <- function(sample, p=0.8, debug=F) {
  table(sample)
  mmx <- seq(min(sample), max(sample))
  pm <- pgeom(mmx, p=p)
  pmf <- pm - c(0, pm[1:length(pm)-1])
  
  e_tbl <- as.vector(table(sample))
  
  tbl <- pmf * sum(e_tbl)
  
  while(length(e_tbl)<length(tbl)) {
    e_tbl <- c(e_tbl, 0)
  }
  
  all_tbl <- rbind(e_tbl, tbl)
  if (debug) { print(all_tbl) }
  
  chi <- chisq.test(e_tbl, tbl, simulate.p.value = TRUE)
  return(chi)
}

bad_test_geom <- function(sample, k=20) {
  interval <- seq(min(sample), max(sample), length.out = k+1)
  empiric <- hist(sample, breaks = interval, plot = F)
  interval[1] <- -Inf
  interval[k+1] <- +Inf
  teor_p <- pgeom(interval, p = 0.8)
  teor_p <- teor_p[2:(k+1)] - teor_p[1:k]
  chi <- chisq.test(empiric$counts, teor_p, simulate.p.value = TRUE)
  return(chi)
}



```
```{r, echo=FALSE}
chi <- test_geom(sample, 0.8, debug=T)
chi
```



```{r, echo=T}
descdist(sample, discrete = T)
```


```{r, echo=F}
fit.geom <- fitdist(sample, distr="geom", method='mle', discrete=T, lower = c(0.1))
plot(fit.geom)
fit.geom
```



***

### Exercise 3

Using Chi-squared test [@Nobody06], we get the following:

```{r}

count_num_bad <- function(N=1000, num_per=200, alpha=0.05) {
  count <- 0
  
  for (i in 1:1000) {
    x <- rgeom(num_per, p=0.8)
    chisq <- test_geom(x, p=0.8)
    if(chisq$p.value < alpha) 
      count <- count + 1
  }
  return(count)
}
N_d=1000
print(paste("Number of 'bad' out of", N_d, "at (0.05) is:", count_num_bad(N=N_d, num_per=200, alpha=0.05)))

```

***

### Задание 4

Имеем следующую функцию:
```{r}

friendly_test_geom <- function(sample, p=0.8, alpha = 0.05){
  chisq <- test_geom(sample, p=p)
  s <- ""
  if (chisq$p.value < alpha){
    s <- " not"
  }
  
  res <- paste0(
    "For significance level alpha=", alpha, 
    ", the sample does", s, " conform to the ", "Geom(0.8)", 
    " distribution."
  )
  return(res)
}
```


```{r, include=FALSE}
print(friendly_test_geom(sample, alpha=0.05))
print(friendly_test_geom(sample, alpha=0.001))
print(friendly_test_geom(sample, alpha=1e-7))
```

***

### Bibliography






